---
title: "Meta plots of tracks for Supercoiling"
author: "Christ Leemans"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
    html_document:
        theme: journal #cerulean
        highlight: monochrome
        toc: true
        toc_float: true
        code_folding: show
    pdf_document:
        fig_width: 8
        fig_height: 5
editor_options:
    chunk_output_type: console
---
# knitr document van Steensel lab

## generated on {date}

In this report I would like to make some nice meta plots for Stephano.
He has discovered regions with opposing effects of TOP1 and DRB treatments.
For these regions I ran a pipeline binning and averaging pA-DamID data and
supercoiling information from the Gilbert lab.



### plots around boundaries.

I took the start and end boundaries of the previously identified regions and
transformed it into a bed file with the start and end boundaries encoded with
opposing orientation.

Using deeptools this will result in a big data matrix with signal for both
boundaries aligned in such a way that the outside of the boundary will
be located downstream of the center while the inside of the boundary will
be located upstream of the boundary.

This way the data can nicely be loaded into R and averaged by replicates.

```{r}
library(data.table)
library(stringr)
library(scales)
library(ggplot2)

file_list = list.files('results_supercoiling/window_runmean',
                       pattern='center.*-550000-550000-5000',
                       full.names=T)
meta.dt = data.table(str_match(file_list, '.*/(.*)_center-(.*)_(.*?)-.*'))

colnames(meta.dt) = c('file', 'region', 'name', 'target')

meta.dt[,rep:=gsub('.*_(.*)', '\\1', name)]
meta.dt[,condition:=gsub('(.*)_.*', '\\1', name)]
meta.dt[,condition:=gsub('_1[0-9]+_', '_', condition)]
meta.dt[,condition:=gsub('_GFP_', '_', condition)]

meta.dt[name=='RPE1', c('rep', 'target', 'name', 'condition') :=
                      list(gsub('replicate', 'r', target),
                           'supercoiling', 'supercoiling',
                           'WT')]

meta.dt[, random:=F]
meta.dt[grep('_random', region), c('region', 'random') :=
                                list(gsub('_random', '', region), T)]



setkey(meta.dt, 'file')
binned.dt = meta.dt[condition%in%c('WT', 'RPE_CT', 'RPE_DRB', 'RPE_SC',
                                   'RPE_Top1'),
                    fread(file, stringsAsFactors=F, header=T), by=file]

binned_meta = merge(binned.dt, meta.dt)
setkey(binned_meta, 'transcript_ID')


pos_columns = colnames(binned.dt[, -c('file', 'transcript_ID')])

up_down = binned_meta[target=='supercoiling',
                       list(left = mean(unlist(.SD[, 1:(ncol(.SD)/2)]), na.rm=T),
                            right = mean(unlist(.SD[, -(1:(ncol(.SD)/2))]), na.rm=T)),
                       by=c('transcript_ID', 'region', 'random'),
                       .SDcols=pos_columns]

binned_updown = merge(binned_meta, up_down, by=c('region', 'random', 'transcript_ID'))



flip <- function(dt, left, right){
    if (left > right){
        return(dt)
    } else {
        invers = dt[,ncol(dt):1]
        colnames(invers) = colnames(dt)
        return(invers)
    }
}

flipped.dt = binned_updown[, flip(.SD, left, right),
                           by=c('file', 'transcript_ID'), .SDcols=pos_columns]

file_mean = flipped.dt[, list(pos = as.numeric(colnames(.SD)),
                              log2_mean = colMeans(log2(.SD))), by='file',
                              .SDcols=-'transcript_ID']

file_mean_merge = merge(file_mean, meta.dt, by='file')


mean_rep.dt = file_mean_merge[, list(log2_mean=mean(log2_mean)),
                               by=c('region', 'target', 'condition', 'random',
                                    'pos') ]

mean_rep.dt[,random:=ifelse(random==T, 'Randomized Regions', 'Actual Regions')]

setkeyv(mean_rep.dt, c('region', 'target', 'random'))


plot_list = apply(unique(mean_rep.dt[,c('region', 'target', 'random')]),
                  1, function(x){
                      dt = mean_rep.dt[as.list(x), ]
                      ggplot(dt, aes(x=pos, y=log2_mean, color=condition)) +
                          geom_line() +
                          ggtitle(paste(x, collapse=' :: ')) +
                          theme_bw() +
                          scale_x_continuous(lim=c(-500000, 500000), name="Position",
                                             labels = scales::comma)
                  })
pdf('cl20200615_center_profiles.pdf', useDingbats=F)
for (plot in plot_list){
    print(plot)
}
dev.off()

setkeyv(mean_rep.dt, c('region', 'target'))

plot_list = apply(unique(mean_rep.dt[target=='supercoiling',c('region', 'target')]),
                  1, function(x){
                      dt = mean_rep.dt[as.list(x), ]
                      ggplot(dt, aes(x=pos, y=log2_mean, color=random)) +
                          geom_line() +
                          ggtitle(paste(x, collapse=' :: ')) +
                          theme_bw() +
                          scale_x_continuous(lim=c(-500000, 500000), name="Position",
                                             labels = scales::comma)
                  })
pdf('cl20200615_center_profiles_supercoiling only.pdf', useDingbats=F)
for (plot in plot_list){
    print(plot)
}
dev.off()


plot_list = apply(unique(mean_rep.dt[,c('region', 'target')]),
                  1, function(x){
                      dt = mean_rep.dt[as.list(x), ]
                      ggplot(dt, aes(x=pos, y=log2_mean, color=condition,
                                     linetype=random)) +
                          geom_line() +
                          ggtitle(paste(x, collapse=' :: ')) +
                          theme_bw() +
                          scale_x_continuous(lim=c(-500000, 500000), name="Position",
                                             labels = scales::comma)
                  })
pdf('cl20200615_center_profiles_random_same_plot.pdf', useDingbats=F)
for (plot in plot_list){
    print(plot)
}
dev.off()
```


```{r}

file_list = list.files('results_supercoiling/scaled_runmean',
                       full.names=T)
meta.dt = data.table(str_match(file_list, '.*/(.*?)-(.*)_(.*?)[.].*'))

colnames(meta.dt) = c('file', 'region', 'name', 'target')

meta.dt[,rep:=gsub('.*_(.*)', '\\1', name)]
meta.dt[,condition:=gsub('(.*)_.*', '\\1', name)]
meta.dt[,condition:=gsub('_1[0-9]+_', '_', condition)]
meta.dt[,condition:=gsub('_GFP_', '_', condition)]

meta.dt[name=='RPE1', c('rep', 'target', 'name', 'condition') :=
                      list(gsub('replicate', 'r', target),
                           'supercoiling', 'supercoiling',
                           'WT')]

meta.dt[, random:=F]
meta.dt[grep('_random', region), c('region', 'random') :=
                                 list(gsub('_random', '', region), T)]


setkey(meta.dt, 'file')
binned.dt = meta.dt[condition%in%c('WT', 'RPE_CT', 'RPE_DRB', 'RPE_SC',
                                   'RPE_Top1'),
                    fread(file, stringsAsFactors=F, header=T), by=file]

file_mean = binned.dt[, list(pos = factor(colnames(.SD), levels=colnames(.SD)),
                             log2_mean = colMeans(log2(.SD))), by='file',
                      .SDcols=-'transcript_ID']

file_mean_merge = merge(file_mean, meta.dt, by='file')

mean_rep.dt = file_mean_merge[random==F, list(log2_mean=mean(log2_mean)),
                               by=c('region', 'target', 'condition', 'pos')]

setkeyv(mean_rep.dt, c('region', 'target'))

pos_levels = mean_rep.dt[,levels(pos)]

xlim = c(which(pos_levels=='TSS-500000'), which(pos_levels=='TES+500000'))

break_vec = c(which(pos_levels=='TSS-500000'),
              which(pos_levels=='TSS-250000'),
              which(pos_levels=='TSS+0'),
              which(pos_levels=='MID+0'),
              which(pos_levels=='MID+500000'),
              which(pos_levels=='TES-500000'),
              which(pos_levels=='TES+0'),
              which(pos_levels=='TES+250000'),
              which(pos_levels=='TES+500000'))

break_names = c('-500kb', '-250kb', 'Edge', 'Edge+250kb', 'scaled-center',
                'Edge-250kb', 'Edge', '+250kb', '+500kb')
plot_list = apply(unique(mean_rep.dt[,c('region', 'target')]),
                  1, function(x){
                      dt = mean_rep.dt[as.list(x), ]
                      ggplot(dt, aes(x=as.numeric(pos), y=log2_mean, color=condition)) +
                          geom_line() +
                          ggtitle(paste(x, collapse=' :: ')) +
                          theme_bw() +
                          scale_x_continuous(breaks=break_vec, lim=xlim,
                                             labels=break_names)
                  })
pdf('cl20200609_region_profiles.pdf', useDingbats=F, width=20)
for (plot in plot_list){
    print(plot)
}
dev.off()
```



```{r}
mid_columns = grep('MID', colnames(binned.dt), value=T)

do_mean <- function(x){
    return(mean(x))
}

setkey(binned.dt, 'transcript_ID')

binned_meta = merge(binned.dt, meta.dt)

up_down = binned_meta[target=='supercoiling',
                       list(left = mean(unlist(.SD[, 1:(ncol(.SD)/2)]), na.rm=T),
                            right = mean(unlist(.SD[, -(1:(ncol(.SD)/2))]), na.rm=T)),
                       by=c('transcript_ID', 'region', 'random'),
                       .SDcols=mid_columns]

binned_updown = merge(binned_meta, up_down, by=c('region', 'random', 'transcript_ID'))

pos_columns = colnames(binned.dt[, -c('file', 'transcript_ID')])


flip <- function(dt, left, right){
    if (left > right){
        return(dt)
    } else {
        invers = dt[,ncol(dt):1]
        colnames(invers) = colnames(dt)
        return(invers)
    }
}

flipped.dt = binned_updown[, flip(.SD, left, right),
                           by=c('file', 'transcript_ID'), .SDcols=pos_columns]

file_mean = flipped.dt[, list(pos = factor(colnames(.SD), levels=colnames(.SD)),
                              log2_mean = colMeans(log2(.SD))), by='file',
                              .SDcols=-'transcript_ID']

file_mean_merge = merge(file_mean, meta.dt, by='file')

mean_rep.dt = file_mean_merge[, list(log2_mean=mean(log2_mean)),
                              by=c('region', 'target', 'condition', 'random', 'pos')]

setkeyv(mean_rep.dt, c('region', 'target', 'random'))

pos_levels = mean_rep.dt[,levels(pos)]

xlim = c(which(pos_levels=='TSS-500000'), which(pos_levels=='TES+500000'))

break_vec = c(which(pos_levels=='TSS-500000'),
             which(pos_levels=='TSS-250000'),
             which(pos_levels=='TSS+0'),
             which(pos_levels=='MID+0'),
             which(pos_levels=='MID+500000'),
             which(pos_levels=='TES-500000'),
             which(pos_levels=='TES+0'),
             which(pos_levels=='TES+250000'),
             which(pos_levels=='TES+500000'))

break_names = c('-500kb', '-250kb', 'Edge', 'Edge+250kb', 'scaled-center',
               'Edge-250kb', 'Edge', '+250kb', '+500kb')
plot_list = apply(unique(mean_rep.dt[,c('region', 'target', 'random')]),
                 1, function(x){
                     x_list = as.list(x)
                     x_list$random = ifelse(x_list[['random']]=='FALSE', F, T)
                     dt = mean_rep.dt[x_list, ]
                     title = paste(x[names(x_list)!='random'], collapse=' :: ')
                     if (x_list$random==T){
                         title = paste('Randomly drawn based on', title)
                     }
                     ggplot(dt, aes(x=as.numeric(pos), y=log2_mean, color=condition)) +
                         geom_line() +
                         ggtitle(title) +
                         theme_bw() +
                         scale_x_continuous(breaks=break_vec, lim=xlim,
                                            labels=break_names)
                 })
pdf('cl20200609_region_profiles_flipped_SuperCoil.pdf', useDingbats=F, width=20)
for (plot in plot_list){
   print(plot)
}
dev.off()
```
