---
title: "Meta plots of DamID on DRB gain/loss and TOP1 gain/loss"
author: "Christ Leemans"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
    html_document:
        theme: journal #cerulean
        highlight: monochrome
        toc: true
        toc_float: true
        code_folding: show
    pdf_document:
        fig_width: 8
        fig_height: 5
editor_options:
    chunk_output_type: console
---
# knitr document van Steensel lab

## generated on {date}

In this report I would like to make some nice meta plots for Stephano.
He has discovered regions with opposing effects of TOP1 and DRB treatments.
For these regions I ran a pipeline binning and averaging pA-DamID.



### plots on TSS.

I used deeptools to calculate the DamID signal 40kb downstream and 20kb upstream of the TSS. I used 200bp bins and 20 bin smooth.

```{r}
library(data.table)
library(stringr)
library(scales)
library(ggplot2)

file_list = list.files('cl20200811_TSS_window_TOP_DRB/window_runmean',
                       pattern='gencode.v24.*100000-100000-1000', full.names=T)
meta.dt = data.table(str_match(file_list, '.*/(.*?)-(.*)_(.*?)-.*'))

colnames(meta.dt) = c('file', 'region', 'name', 'target')

meta.dt[,rep:=gsub('.*_(.*)', '\\1', name)]
meta.dt[,condition:=gsub('(.*)_.*', '\\1', name)]
meta.dt[,condition:=gsub('_1[0-9]+_', '_', condition)]
meta.dt[,condition:=gsub('_GFP_', '_', condition)]

meta.dt[name=='RPE1', c('rep', 'target', 'name', 'condition') :=
                      list(gsub('replicate', 'r', target),
                           'supercoiling', 'supercoiling',
                           'WT')]


setkey(meta.dt, 'file')
binned.dt = meta.dt[condition%in%c('RPE_CT', 'RPE_DRB-3h', 'RPE_TPL-3h') &
                    rep %in% c('r2', 'r3') & target=='Lmnb2',
                    fread(file, stringsAsFactors=F, header=T), by=file]

binned_meta = merge(binned.dt, meta.dt)
setkey(binned_meta, 'transcript_ID')


delta_genes = data.table(readRDS('/DATA/usr/s.manzo/Projects/Christ/genes_Delta_LB2_genes_All.rds'))

transcript_id = fread('cl20200811_TSS_window_TOP_DRB/selection/gencode.v24-fantom-selection.txt',
                      col.names=c('seqnames', 'tss', 'transcript_ID', 'strand',
                                  'gene_id', 'avg', 'tissue_count'),
                      key='transcript_ID')

binned_meta_id = merge(transcript_id[,c('transcript_ID','gene_id')], binned_meta)

binned_delta = merge(binned_meta_id, delta_genes, by='gene_id')

pos_columns = colnames(binned.dt[, -c('file', 'transcript_ID')])


region_list = lapply(c('D_DRB3h', 'D_TPL3h'), function(exp){
    reg_list = lapply(c(-0.3, -0.2, 0.2, 0.3), function(delta){
        if (delta < 0){
            selection = paste(exp, '<', delta)
        } else {
            selection = paste(exp, '>', delta)
        }
        dt = binned_delta[eval(parse(text=paste(selection))) &
                          size < 100000,
                                list(pos = as.numeric(colnames(.SD)),
                                     log2_mean = colMeans(log2(.SD), na.rm=T),
                                     region = paste(exp, delta, sep='_')),
                                by=c('file', 'condition', 'target'),
                                .SDcols=pos_columns]
        return(dt)
    })
    dt = do.call(rbind, reg_list)
    return(dt)
})
region_dt = do.call(rbind, region_list)

mean_rep.dt = region_dt[, list(log2_mean=mean(log2_mean, na.rm=T)),
                        by=c('region', 'target', 'condition', 'pos') ]


setkeyv(mean_rep.dt, c('region', 'target'))


plot_list = apply(unique(mean_rep.dt[,c('region', 'target')]),
                  1, function(x){
                      print(as.list(x))
                      dt = mean_rep.dt[as.list(x), ]
                      ggplot(dt, aes(x=pos, y=log2_mean, color=condition)) +
                          geom_line() +
                          ggtitle(paste(x, collapse=' :: ')) +
                          theme_bw()
                  })
pdf('cl20200819_TSS_profiles_DRB_TPL.pdf', useDingbats=F)
for (plot in plot_list){
    print(plot)
}
dev.off()


```



```{r}

region_list = lapply(c('D_DRB3h', 'D_TPL3h'), function(exp){
    reg_list = lapply(c(-0.3, -0.2, 0.2, 0.3), function(delta){
        if (delta < 0){
            selection = paste(exp, '<', delta)
        } else {
            selection = paste(exp, '>', delta)
        }
        dt = binned_delta[eval(parse(text=paste(selection))) &
                          (CT_All > 1 | DRB_All > 1) &
                          size < 100000,
                                list(pos = as.numeric(colnames(.SD)),
                                     log2_mean = colMeans(log2(.SD), na.rm=T),
                                     region = paste(exp, delta, sep='_')),
                                by=c('file', 'condition', 'target'),
                                .SDcols=pos_columns]
        return(dt)
    })
    dt = do.call(rbind, reg_list)
    return(dt)
})
region_dt = do.call(rbind, region_list)

mean_rep.dt = region_dt[, list(log2_mean=mean(log2_mean, na.rm=T)),
                        by=c('region', 'target', 'condition', 'pos') ]


setkeyv(mean_rep.dt, c('region', 'target'))


plot_list = apply(unique(mean_rep.dt[,c('region', 'target')]),
                  1, function(x){
                      print(as.list(x))
                      dt = mean_rep.dt[as.list(x), ]
                      ggplot(dt, aes(x=pos, y=log2_mean, color=condition)) +
                          geom_line() +
                          ggtitle(paste(x, collapse=' :: ')) +
                          theme_bw()
                  })
pdf('cl20200819_TSS_profiles_DRB_TPL_LAD.pdf', useDingbats=F)
for (plot in plot_list){
    print(plot)
}
dev.off()

```
